Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-05-28 manuscript/manuscript.Rmd
updates MCMC setion with STAN talk

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms except climate covariates were fit by species, while the climate covariates were fit hierarchically with species-specific coefficients drawn from a shared `global' coefficient distribution. We did so for two reasons: (1) the four focal species are all perennial grasses that we expect to respond similarly to climate covariates, and (2) convergence of climate effects at the quadrat level was much easier to achieve when we modeled these terms hierarchically, allowing them to "share" statistical strength via partial pooling [@Gelman2007]. So, climate effects were modeled as:

\begin{align}
\theta_{jk} &\sim \text{Normal}(\bar{\theta_{k}}, \sigma_{k})
\end{align}

where $\bar{\theta_{k}}$ is the interspecific effect of the $k$th climate covariate.

We used uninformative priors for all unknown parameters, specifically:

\begin{align}
\boldsymbol{\gamma, \beta, \bar{\theta}} &\sim \text{Normal}(0, 1e^{-6}) \\
\boldsymbol{\phi} &\sim \text{Normal}(0, \boldsymbol{\sigma_{\phi}}) \\
\boldsymbol{\sigma_{\phi}} &\sim e^{(\text{Gamma}(2, 0.5))} \\
\boldsymbol{\sigma_{\theta}, \sigma_{\gamma}, \sigma_{\beta}, \tau, \zeta} &\sim \text{Gamma}(0.001, 0.001)
\end{align}

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the MCMC sampler in JAGS [@Plummer2003] to estimate the posterior distributions of model parameters and the package 'rjags' to connect R to JAGS. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, thinned by 50, after discarding an initial 50,000 iterations. We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.01, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation.

>>>>

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. We used flat, uninformative priors for all unknown parameters (Appendix X).

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the `No-U-Turn` MCMC sampler in STAN to estimate the posterior distributions of model parameters using the package 'rstan'. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise some more familiar with other MCMC samplers.
We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.01, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation.
